---
title: "Blue Apron"
author: "Andrey Kasyanov"
date: "December 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=1, digits=8)
```

```{r, eval=TRUE, warning=FALSE}
library(data.table)
library(tidyverse)
library(lfe)
library(ggplot2)
library(stargazer)
```

* Experimental design:
Diff in diff experiment
We measure profit-defining factors of 3 periods: pre treatment, during treatment and after treatment, 2 months for each period.
Locations: Boston / Chicago
Duration: 2 + 2 + 2 months
Treatment: Adding or similar priced vegetarian friendly-items in addition to meal kits in boxes of treatment group. Boston is treatment, Chicago is control
Outcome variables: Revenue and Profit

```{r, eval=TRUE, warning=FALSE}
num_obj <- 1000
boston <- rbinom(num_obj, 1, prob = .5)
#population samples for boston and chicago
num_bos <- sum(boston)
num_chi <- num_obj - num_bos
# Boston: Treatment Group / Chicago: Control Group
free_salad <- boston
```

Blue apron offers two meal plans, family and 2-person. Though we have not been able to get exact distribution, BA's revenue statements suggests most people stick with 2 person plan, as on average a customer brings in 250$/month revenue, which is a month's worth of 2 person plan ($59.94/week), compared to family's plan ($71.92/week)

```{r, eval=TRUE, warning=FALSE}
# cost function
randomCost <- function(n, m, sd) rnorm(n, m, sd)
randomError <- function(n, m, sd) rnorm(n, m, sd)

#treatment and control populations 
pre_treat_num <- rnorm(num_bos, 3, 1)
pre_control_num <- rnorm(num_chi, 3, 1)

#revenue is fixed for both control and treatment as it is a price per box
pre_treat_revenue <- pre_treat_num * 60
pre_control_revenue <- pre_control_num * 60

# cost
cost_mean <- 43
cost_sd <- 2
pre_cost <- randomCost(num_obj, cost_mean, cost_sd) + randomError(num_obj, 0, 1)
#profit is not fixed, as it includes the cost of including a salad

#treat_profit <- treat_num * #find profit
#control_profit <- treat_num * #find profit -3

t <- 0
c <- 0
pre_revenue <- vector()
pre_profit <- vector()

# merge since we have 3 periods, this function maps pre treatment values for control and treatment values.
for(i in boston){
  if(i == 1) {
    t <- t + 1
    pre_revenue <- c(pre_revenue, pre_treat_revenue[t])
    pre_profit <- c(pre_profit, pre_treat_revenue[t] - (pre_treat_num[t] * pre_cost[t+c]))
  } else {
    c <- c + 1
    pre_revenue <- c(pre_revenue, pre_control_revenue[c])
    pre_profit <- c(pre_profit, pre_control_revenue[c] - (pre_control_num[c] * pre_cost[t+c]))
  }
}
```

```{r, eval=TRUE, warning=FALSE}
#treatment and control populations 
during_treat_num <- rnorm(num_bos, 6, 1)
during_control_num <- rnorm(num_chi, 3, 1)

#revenue is fixed for both control and treatment as it is a price per box
during_treat_revenue <- during_treat_num * 60
during_control_revenue <- during_control_num * 60

# cost
# We have calculated approximate profit margin per box per person to be approximately 25%. Therefore the cost is around 75%, around $43
cost_mean <- 43
cost_sd <- 3
during_cost <- randomCost(num_obj, cost_mean, cost_sd) + randomError(num_bos, 0, 1)
salad_cost <- 3


t <- 0
c <- 0
during_revenue <- vector()
during_profit <- vector()

#This function maps treatment time profit for control and treatment populations.

for(i in boston){
  if(i == 1) {
    t <- t + 1
    during_revenue <- c(during_revenue, during_treat_revenue[t])
    during_profit <- c(during_profit, during_treat_revenue[t] - (during_treat_num[t] * during_cost[t+c]) - (during_treat_num[t] * salad_cost))
  } else {
    c <- c + 1
    during_revenue <- c(during_revenue, during_control_revenue[c])
    during_profit <- c(during_profit, during_control_revenue[c] - (during_control_num[c] * during_cost[t+c]))
  }
}
```

```{r, eval=TRUE, warning=FALSE}
#treatment and control populations 
after_treat_num <- rnorm(num_bos, 4, 1.5)
after_control_num <- rnorm(num_chi, 3, 1)

#revenue is fixed for both control and treatment as it is a price per box
after_treat_revenue <- after_treat_num * 60
after_control_revenue <- after_control_num * 60

# cost
cost_mean <- 43
cost_sd <- 2
after_cost <- randomCost(num_obj, cost_mean, cost_sd) + randomError(num_bos, 0, 1)
#profit is not fixed, as it includes the cost of including a salad

#treat_profit <- treat_num * #find profit
#control_profit <- treat_num * #find profit -3

t <- 0
c <- 0
after_revenue <- vector()
after_profit <- vector()
# This function maps post treatment values for control and treatment populations. 
for(i in boston){
  if(i == 1) {
    t <- t + 1
    after_revenue <- c(after_revenue, after_treat_revenue[t])
    after_profit <- c(after_profit, after_treat_revenue[t] - (after_treat_num[t] * after_cost[t+c]))
  } else {
    c <- c + 1
    after_revenue <- c(after_revenue, after_control_revenue[c])
    after_profit <- c(after_profit, after_control_revenue[c] - (after_control_num[c] * after_cost[t+c]))
  }
}
```

```{r, eval=TRUE, warning=FALSE}
# covariates
# veg
#According to national us census, vegetarian's occupy around 6% of population, so it is a valid assumption that this distribution is similar in large random sample.
#BA has same pricing for vegetarian and non vegetarian options, so the mean rvenue does not change much. However as it is generally more expensive and taxing to be #vegetarian when it comes to dietary requirement and compensating for lack of animal products, we believe that BA's boxes might present an easier alternative for them.

veg <- rbinom(num_obj, 1, prob = .06)

#referred by a friend
#we found that approximately 34% of blue apron's new customers were acquired by the word of mouth (referals), hence it is worth testing for the impact of including free salad on referals
#Since one person can refer multiple people and get a discount, we use a dummy variable for it. This helps us to avoid complications associated with discounts from referals, which apply if a person refered more than one of their acquaintances.
pre_ref <- rbinom(num_obj, 1, prob = .34)
during_ref <- rbinom(num_obj, 1, prob = .34)
after_ref <- rbinom(num_obj, 1, prob = .34)

# age
# We used Blue Apron's customers' age data for our distribution
rollAge <- function(n) sample(1:100,n,replace=T)

ageDist <- rollAge(num_obj)

age <- vector()
#for our regression we plan using 6 age groups to see which groups has more impact on the profits.
for (j in ageDist) {
  if (j < 4) {
    age <- c(age, '18-24')
  } else if (j < 40) {
    age <- c(age, '25-34')
  } else if (j < 65) {
    age <- c(age, '35-44')
  } else if (j < 82) {
    age <- c(age, '45-54')
  } else if (j < 95) {
    age <- c(age, '55-64')
  } else {
    age <- c(age, '65-99')
  } 
}



#summary table 
blue_apron <- data.table(
  id = 1:num_obj,
  boston = boston,
  free_salad = free_salad,
  pre_revenue = pre_revenue,
  pre_profit = pre_profit,
  during_revenue = during_revenue,
  during_profit = during_profit,
  after_revenue = after_revenue,
  after_profit = after_profit,
  veg = veg,
  pre_ref = pre_ref,
  during_ref = during_ref,
  after_ref = after_ref,
  age = age
)

blue_apron
```

```{r, eval=TRUE, warning=FALSE}
period <- vector()
profit <- vector()
revenue <- vector()
treat <- vector()
ref <- vector()
#regressions and analysis
#this allows us to separate variables into True and False, originally we had encountered an issue of the graph displaying treatment and controls as continious values between 0 and 1, even though we only had 0 and 1, nothing in between.
for(i in free_salad){
  if(i==1){
    treat <- c(treat, "TRUE")
  } else {
    treat <- c(treat, "FALSE")
  }
}
#this function goes through the data set, mapping profit on a graph.
for(i in pre_profit){
  period <- c(period, 0)
  profit <- c(profit, log10(i))
}
for(i in during_profit){
  period <- c(period, 1)
  profit <- c(profit, log10(i))
}
for(i in after_profit){
  period <- c(period, 2)
  profit <- c(profit, log10(i))
}

for(i in pre_revenue){
  revenue <- c(revenue, log10(i))
}
for(i in during_revenue){
  revenue <- c(revenue, log10(i))
}
for(i in after_revenue){
  revenue <- c(revenue, log10(i))
}

for(i in pre_ref){
  ref <- c(ref, i)
}
for(i in during_ref){
  ref <- c(ref, i)
}
for(i in after_ref){
  ref <- c(ref, i)
}


dt <- cbind(period, profit)
dt <- cbind(dt, revenue)
dt <- cbind(dt, ref)
dt <- cbind(dt, treat)
dt <- data.table(dt)
colnames(dt) <- c("period", "profit", "revenue", "referring", "treat")
dt
```
```{r}
#graph on the value of profitabilty and influences of free salad inclusion.
aprn_diff_profit_plot <- ggplot(dt, aes(x=period, y=as.numeric(profit), group = treat)) + geom_smooth(aes(color=treat)) + geom_point(aes(color=treat)) + labs(y="profit (log10)")
aprn_diff_profit_plot
```

```{r}
#graph on the value of revenue and influences of free salad inclusion.
aprn_diff_rev_plot <- ggplot(dt, aes(x=period, y=as.numeric(revenue), group = treat)) + geom_smooth(aes(color=treat)) + geom_point(aes(color=treat)) + labs(y="revenue (log10)")
aprn_diff_rev_plot
```

```{r}
#regression on the value of profitabilty and influences of free salad inclusion with age as covariate, diet as fixed effects
during_pre_profit <- felm(during_profit-pre_profit ~ free_salad + age | veg, blue_apron)
after_pre_profit <- felm(after_profit-pre_profit ~ free_salad + age | veg, blue_apron)
after_during_profit <- felm(after_profit-during_profit ~ free_salad + age | veg, blue_apron)
stargazer(during_pre_profit, after_pre_profit, after_during_profit, type = 'text')
```

```{r}
#regression on the value of profitabilty and influences of free salad inclusion with diet as covariate, age as fixed effect
during_pre_profit <- felm(during_profit-pre_profit ~ free_salad + veg | age, blue_apron)
after_pre_profit <- felm(after_profit-pre_profit ~ free_salad + veg | age, blue_apron)
after_during_profit <- felm(after_profit-during_profit ~ free_salad + veg | age, blue_apron)
stargazer(during_pre_profit, after_pre_profit, after_during_profit, type = 'text')
```




```{r}
#regression on the value of revenue and influences of free salad inclusion with age as covariate, diet as fixed effect
during_pre_revenue <- felm(during_revenue-pre_revenue ~ free_salad + age | veg, blue_apron)
after_pre_revenue <- felm(after_revenue-pre_revenue ~ free_salad + age | veg, blue_apron)
after_during_revenue <- felm(after_revenue-during_revenue ~ free_salad + age | veg, blue_apron)
stargazer(during_pre_revenue, after_pre_revenue, after_during_revenue, type = 'text')
```
```{r}
#regression on the number of referrals and influences of free salad inclusion with age as covariate, die as fixed effect
after_pre_referred <- felm(after_ref-pre_ref ~ free_salad + age | veg, blue_apron)
after_during_referred <- felm(after_ref-during_ref ~ free_salad + age | veg, blue_apron)
during_pre_referred <- felm(during_ref-pre_ref ~ free_salad + age | veg, blue_apron)

stargazer(during_pre_referred, after_pre_referred, after_during_referred, type = 'text')
```
```{r}
#balance test to ensure parallel trends
balance_test <- felm(pre_profit ~ free_salad, data = blue_apron)
summary(balance_test)
```
